---
title: "Assignment 2"
author: "Tarun Sudhams"
date: "11/12/2018"
output: html_document
---

# Question 1 (Chapter 6, Exercise 8 (Page 262))
<br>
*In this exercise, we will generate simulated data, and will then use this data to perform best subset selection.*

*(a) Use the rnorm() function to generate a predictor X of length n = 100, as well as a noise vector eps of length n = 100.*

```{r}
set.seed(3876)
x<-rnorm(100)
error<-rnorm(100)
```
<br>
*(b) Generate a response vector Y of length n = 100 according to the model Y = β0 + β1X + β2X2 + β3X3 + error, where β0, β1, β2, and β3 are constants of your choice.*

```{r}
b0 <- 2
b1 <- 3
b2 <- -1
b3 <- 0.5
y <- b0 + b1 * x + b2 * x^2 + b3 * x^3 + error
```
<br>
*(c) Use the regsubsets() function to perform best subset selection in order to choose the best model containing the predictors X,X2, . . .,X10. What is the best model obtained according to Cp, BIC, and adjusted R2? Show some plots to provide evidence for your answer, and report the coefficients of the best model obtained.Note you will need to use the data.frame() function to create a single data set containing both X and Y.*

```{r}
library(leaps)
data.full<-data.frame(y=y,x=x)
regfit.bwd<-regsubsets(y~x+I(x^2) + I(x^3) + I(x^4) + I(x^5) + I(x^6) + I(x^7) + I(x^8) + I(x^9) + I(x^10), data = data.full, nvmax = 10, really.big = T, method = "backward")
reg.summary.bwd<-summary(regfit.bwd)
par(mfrow=c(1,3))

#Plotting Model with C_P value 
plot(reg.summary.bwd$cp,xlab="Number of Variables", ylab="C_P", type="1")
```
<br>

